---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(egg)
```
# dftplotR

dftplotR is an R package to provide standardised charts formatting in ggplot2. It includes theme functions which meet Government Statistical Service (GSS) best practice guidance. It also includes colour palettes based on DfT corporate branding which meet WCAG 2.0 accessibility standards, and are distinguishable in greyscale.

## Installation

You can install dftplotR with:

```
install.packages("devtools")
devtools::install_github("departmentfortransport/dftplotr")
```

## Overview

The package contains the following functions:

```scale_colour_dft```: ggplot function which applies one of a range of DfT palettes to a ggplot line chart (or other chart which groups by **colour**)

```scale_fill_dft```: ggplot function which applies one of a range of DfT palettes to a ggplot line chart (or other chart which groups by **fill**)

```display_palette```: a visual output which shows all of the colours contained in the selected DfT palette

```palette_picker_tool```: loads an interactive Shiny dashboard which allows users to build their own custom palettes of up to 5 colours using the DfT corporate colours.


## Colour palettes

The dftplotR package comes with 6 pre-loaded palettes. These contain colours taken from the DfT corporate colour scheme, and have also been selected to meet accessibility guidance. The two primary palettes are `main.palette`; a desaturated palette ideal for use in bar charts or other area fill charts, and `electric.brights`, a saturated palette which improves visibility in line-charts or other point colour charts. The other palettes are ideal where these palettes are not appropriate, or where variation is needed.

The palettes and the colours in them can be seen here, or visualised within the package by calling `display.palette()` and the palette name.

```{r, out.width = "80%", fig.align = 'center'}
egg::ggarrange(dftplotr::display_palette("main.palette"),
               dftplotr::display_palette("electric.brights"),
               dftplotr::display_palette("just.beachy"),
               dftplotr::display_palette("cycling.hills"),
               dftplotr::display_palette("mountain.train"),
               dftplotr::display_palette("clear.skies"),
              ncol = 3
)
               
```


## Gradient palettes

The dftplotR package also comes with 6 gradient palettes. These have defined start and end colours taken from the DfT corporate colour scheme, and will auto-generate shades between these two points to create the specified number of colours. These do not meet accessibility guidance, and should only be used when the pre-set palettes are not appropriate and colour is not conveying the primary message of the visualisation.

The palettes and the colours in them can be seen here, or visualised within the package by calling `display.palette(gradient = TRUE)` and the palette name. The number of shades in the palette can be specified using the `n = ` argument, e.g. `n = 8` for a palette with 8 shades. When this argument is not used, the default number of shades is 5. Example gradient palettes with 5 shades can be seen below:

```{r, out.width = "80%", fig.align = 'center', warning=FALSE}
egg::ggarrange(dftplotr::display_palette("main.palette", gradient = TRUE),
               dftplotr::display_palette("electric.brights", gradient = TRUE),
               dftplotr::display_palette("just.beachy", gradient = TRUE),
               dftplotr::display_palette("cycling.hills", gradient = TRUE),
               dftplotr::display_palette("mountain.train", gradient = TRUE),
               dftplotr::display_palette("clear.skies", gradient = TRUE),
              ncol = 3
)
               
```


## ggplot2 themes

## Palette picker tool

## Accessibility

The mojchart colour palettes aim to be accessible to those with the most common forms of colour blindness, and should also be discernable in greyscale. The standard palettes in this package have been designed to meet WCAG 2.0 accessibility guidance.

Four of the created palettes (`main.palette`, `electric.brights`, `just.beachy` and `cycling.hills`) contain four colours each, and meet WCAG 2.0 AAA guidance. All colours have a contrast ratio of at least 4.5:1 to the bars adjacent to them, and alternate light and dark shades to further increase contrast. This is the highest standard of accessibility and should be used when possible.

Two of the created palettes (`mountain.train` and `clear.skies`) contain 6 colours each, and meet WCAG 2.0 AA guidance. All colours have a contrast ration of at least 3:1 to the bars adjacent to them, and alternate light and dark shades to further increase contrast. This is the minimum standard of accessibility required for publication and use should be limited to times when a larger palette is required.

It is also possible to generate gradient palettes using this package, with colours based on the DfT corporate theme. While these palettes can be generated with any number of shades, there is no guarantee that the contrast between these colours is sufficient to meet accessibility standards. These palettes should only be used in very limited circumstances where colour is not the only means of distinguishing different groupings and none of the standard palettes are suitable.

For line charts, it may be possible that lines with reduced contrast are next to each other. As a result, and in line with GSS recommendation, line charts should be labelled directly where possible rather than using a legend.

